Supplementary Material for `The response of trophic interaction networks to multiple stressors in a marine latitudinal gradient of the Southern Hemisphere'

### 2023-11-21 manuscript.tex
New version to submit to GCB

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[vmargin=1in,hmargin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}

\usepackage{pdflscape,booktabs}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage[running]{lineno}
\linenumbers
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\hypertarget{the-response-of-trophic-interaction-networks-to-multiple-stressors-in-a-marine-latitudinal-gradient-of-the-southern-hemisphere}{%
\section{The response of trophic interaction networks to multiple
stressors in a marine latitudinal gradient of the Southern
Hemisphere}\label{the-response-of-trophic-interaction-networks-to-multiple-stressors-in-a-marine-latitudinal-gradient-of-the-southern-hemisphere}}

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Ecological networks are increasingly used as conceptual frameworks for
  understanding ecological community structure, identifying key species,
  and for ecosystem management and biodiversity conservation. Nowadays,
  there is a growing interest in understanding how ecological networks
  respond to environmental and anthropogenic stressors along
  geographical gradients.
\item
  We review for the first time the state-of-the-art knowledge on
  stressor response of marine food webs along the southwest Atlantic to
  Antarctic gradient (45 - 78ºS), considering the following study areas:
  San Jorge Gulf, Beagle Channel, Burdwood Bank, Scotia Sea, Potter Cove
  (Antarctica) and Weddell Sea (Antarctica). The aim of this review is
  threefold: 1) describe the complexity and structure of marine food
  webs along the southwest Atlantic to Antarctic axis from a network
  perspective; 2) identify the ongoing environmental and anthropogenic
  stressors for each marine ecosystem containing the food webs; and 3)
  resume proven food web changes and elaborate hypotheses on how the
  identified stressors might affect food web features, combining
  information on node- and network-level properties.
\item
  Environmental and anthropogenic stressor effects in the study gradient
  have been mostly assessed individually and at the organism and/or
  population, i.e., at the node level, with one exception: the effect of
  fisheries in the San Jorge Gulf food web. To address the most
  plausible stressors effects on the selected food webs, we built
  hypotheses for each study area considering the following: a)
  stressor(s), b) parameter(s) affected, c) node-level properties of the
  affected species, and d) network-level properties of the food web.
\item
  The most common stressor reported along the gradient is global
  warming, except for Beagle Channel and Burdwood Bank, which are more
  influenced by the introduction of an alien species and fisheries,
  respectively. We provide particular hypotheses on how warming might
  impact food webs.
\item
  We demonstrate the advantages of applying a network approach to better
  understand and predict the effect of stressors in marine ecosystems of
  the Southern Hemisphere.
\end{enumerate}

@@

\hypertarget{introduction}{%
\subsection{1. Introduction}\label{introduction}}

The application of a network perspective has emerged as a powerful tool
to tackle the complexity of species interactions, facilitating a better
understanding of the structure and functioning of ecosystems (Belgrano,
Scharler, Scharler, Dunne, \& Ulanowicz, 2005; Thompson et al., 2012;
D'Alelio, Libralato, Wyatt, \& Ribera d'Alcalà, 2016). Trophic networks
(or food webs) allow identifying properties and key species that may be
crucial for ecosystem stability, and hence important for ecosystem
management and biodiversity conservation (Thompson et al., 2012). There
is a growing interest in understanding how ecological networks respond
to environmental and anthropogenic stressors along geographical
gradients (Baiser, Gotelli, Buckley, Miller, \& Ellison, 2012; Cirtwill,
Stouffer, \& Romanuk, 2015; Strydom et al., 2021; Bauer et al., 2022).
Yet, only a few studies have described variation in food web structure
along latitudinal gradients in marine ecosystems. The few that have come
from the Global North (Wood, Russell, Hanson, Williams, \& Dunne, 2015;
Kortsch et al., 2019; Frelat et al., 2022; Pecuchet et al., 2022),
whereas no studies, nor meta-analyses, on geographical variation in
marine food webs exist for the Global South (Southern Hemisphere).

Here we review for the first time the state-of-the-art knowledge on
stressor response of marine food webs along the southwest Atlantic to
Antarctic gradient (45 - 78ºS, Figure 1). We focus on proven and
expected changes in food webs driven by stressors in selected areas
along this large-scale latitudinal gradient. The aim of this review is
threefold: 1) describe the complexity and structure of marine food webs
along the southwest Atlantic to Antarctic axis from a network
perspective; 2) identify the ongoing environmental and anthropogenic
stressors for each marine ecosystem containing the food webs; and 3)
resume proven food web changes and elaborate hypotheses on how the
identified stressors might affect food web features (e.g.~energy flow,
stability), combining information on node- and network-level properties.
Finally, we suggest which additional data and analyses are needed to
gain insights into the stressors' effects on food web properties in the
southwest Atlantic to Antarctic region.

\hypertarget{the-structure-of-marine-food-webs-in-the-southwest-atlantic---antarctic-region}{%
\subsection{2. The structure of marine food webs in the Southwest
Atlantic - Antarctic
region}\label{the-structure-of-marine-food-webs-in-the-southwest-atlantic---antarctic-region}}

Together, the southwest and the Atlantic sector of the Southern Ocean
comprise one of the most biologically productive regions of the world's
oceans (Acha, Mianzan, Guerrero, Favero, \& Bava, 2004; Whitehouse et
al., 2008; Ballerini et al., 2014; Latorre et al., 2023). The referred
region extends from San Jorge Gulf (45ºS) in the Patagonian shelf to the
Weddell Sea (78ºS) in the Southern Ocean, and covers a well-connected
oceanic latitudinal gradient (R. P. Matano, Palma, \& Piola, 2010;
Steele, Thorpe, \& Turekian, 2010; Guihou, Piola, Palma, \& Chidichimo,
2020).

Throughout this latitudinal gradient, many investigations have been
carried out addressing the trophic ecology of specific species and
prey-predator relationships (Vinuesa \& Varisco, 2007; Sailley et al.,
2013; Pasotti et al., 2015; Riccialdelli, Newsome, Fogel, \& Fernández,
2017; Saunders, Hill, Tarling, \& Murphy, 2019; Riccialdelli et al.,
2020), but few studies have considered the complexity of the ecosystem
in terms of a high resolution of species and their prey-predator
interactions (but see Jacob et al. (2011), Marina et al. (2018), Funes,
Saravia, Cordone, Iribarne, \& Galván (2022), López-López, Genner,
Tarling, Saunders, \& O'Gorman (2022), Rodriguez, Marina, Schloss, \&
Saravia (2022), Marina et al. (in rev.)). Neglecting this complexity
might lead to a misunderstanding of the structure and functioning of the
ecosystems, and ultimately reduce the ability to predict ecosystem
responses to perturbations (Montoya, Woodward, Emmerson, \& Solé, 2009;
Woodward et al., 2010).

@@

San Jorge Gulf is the northernmost study area considered in this review.
It is a partially enclosed basin spanning approximately 34,000 km2 and
\textasciitilde100 m of maximum depths, located from Cabo dos Bahias
(44°55'S) to Cabo tres Puntas (47°06'S) (Figure 1). The shallower
northern and southern ends of the Gulf (with depths \textasciitilde40 m)
present two prominent frontal systems. These systems are the areas of
highest productivity (Bogazzi et al., 2005; Glembocki, Williams,
Góngora, Gagliardini, \& Orensanz, 2015). The Gulf's productivity
supports large invertebrate and vertebrate fisheries (María Eva Góngora,
González-Zevallos, Pettovello, \& Mendía, 2012), as well as marine
mammal and seabird populations (Yorio, 2009). The San Jorge Gulf food
web comprises 165 nodes and 1015 trophic interactions with a connectance
of 0.04. The percentage of top predators is 16\%, 78\% of intermediate
nodes and 6\% of basal nodes; 60\% of predators are omnivorous (Table
1). The most connected nodes are: the Argentine red shrimp Pleoticus
muelleri, the squat lobster Gimothea gregaria, squids (Illex argentinus
as dominant species) and Amphipoda. Notably, these nodes present
mid-trophic positions in the food web (3.0, 2.5, 3.6 and 2,
respectively) (Funes et al., 2022).

@@

The so-called Burdwood Bank ecosystem comprises Marine Protected Areas
Namuncurá - Burdwood Bank I and II, meaning the shallow submarine
plateau named Burdwood Bank with a 200 m isobath boundary, and a deep
slope that reaches 4000 m in depth, respectively (Tombesi, Rabufetti, \&
Lovrich, 2020; Administración de Parques Nacionales, 2022) (Figure 1).
Physical features in the plateau are fairly stable, with salinity
averaging 34 all year round and temperature ranging between 4 and 8ºC
(Acha et al., 2004; Piola \& Falabella, 2009). The plateau is surrounded
by steep flanks of up to 4000 m depth, protected by the Marine Protected
Area Namuncurá - Burdwood Bank II (32,000 km2; \textasciitilde55º -
56ºS, \textasciitilde58º - 62ºW). Intense upwelling and mixing occur in
relation to the slope, entraining deep nutrient-rich waters into the
photic layer (Piola \& Falabella, 2009; Ricardo P. Matano, Palma, \&
Combes, 2019), and resulting in a fairly homogeneous water column both
spatially and temporally (Glorioso \& Flather, 1995; Ricardo P. Matano
et al., 2019). The Burdwood Bank food web comprises 379 nodes and 1788
interactions, with a connectance of 0.01, and an asymmetric degree
distribution (i.e.~most of the species have a relatively low number of
interactions and few species concentrate most of them). Almost half of
the consumers are omnivores (0.48), and the network displays a
small-world pattern (Marina et al., in rev.) (Table 1).

@@

In the Antarctic realm, Potter Cove is a \textasciitilde9 km2 fjord with
a depth range of 0 - 200 m located at 25 de Mayo/King George Island
(62°14'S, 58°38'W, South Shetland Islands) on the West Antarctic
Peninsula (Figure 1). The cove, bordered by the Fourcade Glacier, is
divided into three areas: a) the internal cove, a high
glacier-influenced, soft sediment zone with a 50 m maximum depth; b) the
central cove, a mixed substrate area with low meltwater influence and an
80 m maximum depth; and c) the external cove, ice-free for 60 years with
a 185 m maximum depth and rocky substrate (Jerosch et al., 2018; Neder
et al., 2022; Neder, 2023). Potter Cove's high-latitude location results
in variable environmental conditions due to photoperiod length
seasonality. Sea ice often covers this area in winter (Schloss et al.,
2012). With low phytoplankton biomass, macroalgae, and microphytobenthos
are likely the primary food sources for secondary benthic production (M.
L. Quartino \& Boraso de Zaixso, 2008). The Potter Cove food web
includes 110 nodes and 649 interactions, with an asymmetric degree
distribution, and a connectance value of 0.05 (Table 1). It presents a
modular structure (groups of species interact more strongly with each
other than with species belonging to other groups), that's positively
associated with stability, since perturbation can be retained within
modules (Rodriguez et al., 2022).

Located between 74 and 78ºS, the high Antarctic Weddell Sea shelf spans
approximately 450 km from East to West (Jacob et al., 2011) (Figure 1).
The water depth ranges from 200 to 500 m, with shallower regions being
covered by continental ice that forms the coastline along the eastern
and southern parts of the Weddell Sea. Within this shelf area, exists a
complex three-dimensional benthic habitat characterized by substantial
benthic biomasses and an intermediate to high diversity when compared to
benthic boreal communities (Dayton, 1990; Teixidó, Garrabou, \& Arntz,
2002). The Weddell Sea food web exhibits a high level of network
complexity, featuring the greatest number of nodes (490) and trophic
interactions (16041) among the analyzed food webs in this review (Table
1). Its connectance (0.07) and other structural properties are
intermediate compared to the other food webs. Recently, the interaction
strengths of this food web were estimated, revealing the presence of
numerous weak and few strong interactions, which is consistent with
findings in other complex food webs (McCann, Hastings, \& Huxel, 1998;
Nilsson \& McCann, 2016). Notably, this asymmetric distribution of
interaction strength is likely to promote community persistence.

\footnotesize

Table 1. Complexity and structure properties of the marine food webs
considered in the present review. Refer to Table 3 for definition of
properties. mean TL: mean trophic level. Food webs are ordered by
increasing latitude.

@@

\hypertarget{environmental-and-anthropogenic-stressors-in-the-southwest-atlantic---antarctic-region}{%
\subsection{3. Environmental and anthropogenic stressors in the
Southwest Atlantic - Antarctic
region}\label{environmental-and-anthropogenic-stressors-in-the-southwest-atlantic---antarctic-region}}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[vmargin=1in,hmargin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{pdflscape,booktabs}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage[running]{lineno}
\linenumbers
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\section{The response of trophic interaction networks to multiple
stressors in a marine latitudinal gradient of the Southern
Hemisphere}\label{the-response-of-trophic-interaction-networks-to-multiple-stressors-in-a-marine-latitudinal-gradient-of-the-southern-hemisphere}

Tomás I. Marina\textsuperscript{1}, Leonardo A.
Saravia\textsuperscript{1,2,*}, Iara D. Rodriguez\textsuperscript{3},
Manuela Funes\textsuperscript{4}, Georgina Cordone\textsuperscript{5},
Santiago R. Doyle\textsuperscript{3,6}, Anahí
Silvestro\textsuperscript{6}, David E. Galván\textsuperscript{5},
Susanne Kortsch\textsuperscript{7} \& Fernando Momo\textsuperscript{3,6}

\textsuperscript{1} Centro Austral de Investigaciones Científicas
(CADIC-CONICET), Ushuaia, Argentina;

\textsuperscript{2} Instituto de Ciencias Polares, Ambiente y Recursos
Naturales, Universidad Nacional de Tierra del Fuego (UNTdF), Ushuaia,
Argentina;

\textsuperscript{3} Instituto de Ciencias, Universidad Nacional de
General Sarmiento (UNGS), Los Polvorines, Argentina;

\textsuperscript{4} Instituto de Investigaciones Marinas y Costeras
(IIMyC-CONICET), Mar del Plata, Argentina;

\textsuperscript{5} Centro Para el Estudio de Sistemas Marinos
(CESIMAR-CONICET), Puerto Madryn, Argentina;

\textsuperscript{6} Instituto de Ecología y Desarrollo Sustentable
(INEDES-CONICET-UNLu), Luján, Argentina;

\textsuperscript{7} Tvärminne Zoological Station, University of
Helsinki, Hanko, Finland.

* corresponding author: Leonardo A. Saravia. Centro Austral de
Investigaciones Científicas (CADIC-CONICET), Ushuaia, Argentina.
\href{mailto:lasaravia@untdf.edu.ar}{\nolinkurl{lasaravia@untdf.edu.ar}}

\subsection{Abstract}\label{abstract}

Ecological networks offer valuable insights into community structure,
key species identification, and ecosystem management for biodiversity
conservation. Understanding how these networks react to environmental
and anthropogenic stressors, especially along geographical gradients, is
of increasing interest. This review presents a pioneering analysis of
stressor responses in marine food webs from the southwest Atlantic to
the Antarctic (45 - 78ºS), encompassing areas such as San Jorge Gulf,
Beagle Channel, Burdwood Bank, Scotia Sea, Potter Cove, and the Weddell
Sea in Antarctica. Our objectives are to: 1) describe the structure of
marine food webs along this axis using a network approach; 2) identify
predominant environmental and anthropogenic stressors affecting each
ecosystem; and 3) summarize observed food web changes and hypothesize on
stressor impacts. Our collaborative team, comprising regional experts
and global authorities on high-latitude marine food webs and stressor
effects, ensures a comprehensive and credible literature review. We
assessed the effects of stressors primarily at the species level, with
notable exceptions like fisheries in San Jorge Gulf. Hypotheses for each
study area were formulated considering: a) stressors; b) impacted
parameters; c) node-level species properties; and d) network-level food
web properties. Global warming emerges as the most common stressor
across the gradient, except in the Beagle Channel and Burdwood Bank,
where alien species introduction and fisheries are more influential,
respectively. We offer specific hypotheses on how warming may affect
food webs. Our findings highlight the benefits of a network approach in
understanding and predicting stressor effects in Southern Hemisphere
marine ecosystems. This approach provides a holistic understanding of
ecological networks, enhances our ability to identify key species and
interactions, and offers insights for ecosystem management and
conservation in the face of various stressors.

@@

\subsection{1. Introduction}\label{introduction}

The application of a network perspective has emerged as a powerful tool
to tackle the complexity of species interactions, facilitating a better
understanding of the structure and functioning of ecosystems (Belgrano,
Scharler, Scharler, Dunne, \& Ulanowicz, 2005; Thompson et al., 2012).
Trophic networks (or food webs) allow identifying properties and key
species that may be crucial for ecosystem stability, and hence important
for ecosystem management and biodiversity conservation (Thompson et al.,
2012). There is a growing interest in understanding how ecological
networks respond to environmental and anthropogenic stressors along
geographical gradients (Cirtwill, Stouffer, \& Romanuk, 2015; Bauer et
al., 2022). Yet, only a few studies have described variation in food web
structure along latitudinal gradients in marine ecosystems. The few that
have come from the Global North (Wood, Russell, Hanson, Williams, \&
Dunne, 2015; Kortsch et al., 2019; Pecuchet et al., 2022), whereas no
studies, nor meta-analyses, on geographical variation in marine food
webs exist for the Global South (Southern Hemisphere).

Here we review for the first time the state-of-the-art knowledge on
stressor response of marine food webs along the southwest Atlantic to
Antarctic gradient (45 - 78ºS, Figure 1). We focus on proven and
expected changes in food webs driven by stressors in selected areas
along this large-scale latitudinal gradient. We recruited food web
experts from different marine systems of Argentina and the world (see
co-authors' list). Throughout the year 2023 we maintained regular
discussion meetings, typically held every two or three weeks. The aim of
this review is threefold: 1) describe the complexity and structure of
marine food webs along the southwest Atlantic to Antarctic axis from a
network perspective; 2) identify the ongoing environmental and
anthropogenic stressors for each marine ecosystem containing the food
webs; and 3) resume proven food web changes and elaborate hypotheses on
how the identified stressors might affect food web features (e.g.~energy
flow, stability), combining information on node- and network-level
properties. Finally, we suggest which additional data and analyses are
needed to gain insights into the stressors' effects on food web
properties in the southwest Atlantic to Antarctic region.

\subsection{2. The structure of marine food webs in the Southwest
Atlantic - Antarctic
region}\label{the-structure-of-marine-food-webs-in-the-southwest-atlantic---antarctic-region}

Together, the southwest and the Atlantic sector of the Southern Ocean
comprise one of the most biologically productive regions of the world's
oceans (Acha, Mianzan, Guerrero, Favero, \& Bava, 2004; Latorre et al.,
2023). The referred region extends from San Jorge Gulf (45ºS) in the
Patagonian shelf to the Weddell Sea (78ºS) in the Southern Ocean, and
covers a well-connected oceanic latitudinal gradient (R. P. Matano,
Palma, \& Piola, 2010; Guihou, Piola, Palma, \& Chidichimo, 2020).

Throughout this latitudinal gradient, many investigations have been
carried out addressing the trophic ecology of specific species and
prey-predator relationships (Vinuesa \& Varisco, 2007; Pasotti et al.,
2015; Saunders, Hill, Tarling, \& Murphy, 2019; Riccialdelli et al.,
2020), but few studies have considered the complexity of the ecosystem
in terms of a high resolution of species and their prey-predator
interactions (but see Jacob et al. (2011), Marina et al. (2018), Funes,
Saravia, Cordone, Iribarne, \& Galván (2022), López-López, Genner,
Tarling, Saunders, \& O'Gorman (2022), Rodriguez, Marina, Schloss, \&
Saravia (2022), Marina et al. (in rev.)). Neglecting this complexity
might lead to a misunderstanding of the structure and functioning of the
ecosystems, and ultimately reduce the ability to predict ecosystem
responses to perturbations (Montoya, Woodward, Emmerson, \& Solé, 2009).

@@

San Jorge Gulf is the northernmost study area considered in this review.
It is a partially enclosed basin spanning approximately 34,000 km2 and
\textasciitilde100 m of maximum depths, located from Cabo dos Bahias
(44°55'S) to Cabo tres Puntas (47°06'S) (Figure 1). The shallower
northern and southern ends of the Gulf (with depths \textasciitilde40 m)
present two prominent frontal systems. These systems are the areas of
highest productivity (Glembocki, Williams, Góngora, Gagliardini, \&
Orensanz, 2015). The Gulf's productivity supports large invertebrate and
vertebrate fisheries (Góngora, González-Zevallos, Pettovello, \& Mendía,
2012), as well as marine mammal and seabird populations (Yorio, 2009).
The San Jorge Gulf food web comprises 165 nodes and 1015 trophic
interactions with a connectance of 0.04. The percentage of top predators
is 16\%, 78\% of intermediate nodes and 6\% of basal nodes; 60\% of
predators are omnivorous (Table 1). The most connected nodes are: the
Argentine red shrimp Pleoticus muelleri, the squat lobster Gimothea
gregaria, squids (Illex argentinus as dominant species) and Amphipoda.
Notably, these nodes present mid-trophic positions in the food web (3.0,
2.5, 3.6 and 2, respectively) (Funes et al., 2022).

@@

The so-called Burdwood Bank ecosystem comprises Marine Protected Areas
Namuncurá - Burdwood Bank I and II, meaning the shallow submarine
plateau named Burdwood Bank with a 200 m isobath boundary, and a deep
slope that reaches 4000 m in depth, respectively (Administración de
Parques Nacionales, 2022) (Figure 1). Physical features in the plateau
are fairly stable, with salinity averaging 34 all year round and
temperature ranging between 4 and 8ºC (Acha et al., 2004). The plateau
is surrounded by steep flanks of up to 4000 m depth, protected by the
Marine Protected Area Namuncurá - Burdwood Bank II (32,000 km2;
\textasciitilde55º - 56ºS, \textasciitilde58º - 62ºW). Intense upwelling
and mixing occur in relation to the slope, entraining deep nutrient-rich
waters into the photic layer (Ricardo P. Matano, Palma, \& Combes,
2019), and resulting in a fairly homogeneous water column both spatially
and temporally (Ricardo P. Matano et al., 2019). The Burdwood Bank food
web comprises 379 nodes and 1788 interactions, with a connectance of
0.01, and an asymmetric degree distribution (i.e.~most of the species
have a relatively low number of interactions and few species concentrate
most of them). Almost half of the consumers are omnivores (0.48), and
the network displays a small-world pattern (Marina et al., in rev.)
(Table 1).

@@

In the Antarctic realm, Potter Cove is a \textasciitilde9 km2 fjord with
a depth range of 0 - 200 m located at 25 de Mayo/King George Island
(62°14'S, 58°38'W, South Shetland Islands) on the West Antarctic
Peninsula (Figure 1). The cove, bordered by the Fourcade Glacier, is
divided into three areas: a) the internal cove, a high
glacier-influenced, soft sediment zone with a 50 m maximum depth; b) the
central cove, a mixed substrate area with low meltwater influence and an
80 m maximum depth; and c) the external cove, ice-free for 60 years with
a 185 m maximum depth and rocky substrate (Jerosch et al., 2018). Potter
Cove's high-latitude location results in variable environmental
conditions due to photoperiod length seasonality. Sea ice often covers
this area in winter (Schloss et al., 2012). With low phytoplankton
biomass, macroalgae, and microphytobenthos are likely the primary food
sources for secondary benthic production (M. L. Quartino \& Boraso de
Zaixso, 2008). The Potter Cove food web includes 110 nodes and 649
interactions, with an asymmetric degree distribution, and a connectance
value of 0.05 (Table 1). It presents a modular structure (groups of
species interact more strongly with each other than with species
belonging to other groups), that's positively associated with stability,
since perturbation can be retained within modules (Rodriguez et al.,
2022).

Located between 74 and 78ºS, the high Antarctic Weddell Sea shelf spans
approximately 450 km from East to West (Jacob et al., 2011) (Figure 1).
The water depth ranges from 200 to 500 m, with shallower regions being
covered by continental ice that forms the coastline along the eastern
and southern parts of the Weddell Sea. Within this shelf area, exists a
complex three-dimensional benthic habitat characterized by substantial
benthic biomasses and an intermediate to high diversity when compared to
benthic boreal communities (Teixidó, Garrabou, \& Arntz, 2002). The
Weddell Sea food web exhibits a high level of network complexity,
featuring the greatest number of nodes (490) and trophic interactions
(16041) among the analyzed food webs in this review (Table 1). Its
connectance (0.07) and other structural properties are intermediate
compared to the other food webs. Recently, the interaction strengths of
this food web were estimated, revealing the presence of numerous weak
and few strong interactions, which is consistent with findings in other
complex food webs (Nilsson \& McCann, 2016). Notably, this asymmetric
distribution of interaction strength is likely to promote community
persistence.

\newpage
\footnotesize

Table 1. Complexity and structure properties of the marine food webs
considered in the present review. Refer to Table 2 for definition of
properties. mean TL: mean trophic level. Food webs are ordered by
increasing latitude.

@@

\subsection{3. Environmental and anthropogenic stressors in the
Southwest Atlantic - Antarctic
region}\label{environmental-and-anthropogenic-stressors-in-the-southwest-atlantic---antarctic-region}

A stressor is any environmental or anthropogenic variable that causes a
quantifiable change, irrespective of its direction (increase or
decrease), in a biological response (Orr et al., 2020). In the light of
this, the southwest Atlantic - Antarctic marine biota has been and is
currently subjected to a variety of stressors (e.g.~sea warming, glacial
retreat, ice changes, acidification, species invasion, fisheries, and
contamination). Although it is well-known that multiple stressors act in
concert at any given time (e.g.~warming and fishery; acidification and
contaminants) (Gutt et al., 2021), to date stressor assessments have
been performed individually. Moreover, the potential for interactive
effects of two or more stressors (synergy or antagonism) (Côté, Darling,
\& Brown, 2016) is almost unknown for the region (Rowlands et al.,
2021).

In the following subsections, we describe the main environmental and
anthropogenic stressors, and the species (or trophic species) and
parameters affected reported for the marine ecosystems that contain the
food webs considered in this review. In Table S2 we provide an
exhaustive list of all stressors affecting species inhabiting each area,
considering: the type of stressor, species and parameter affected, and
locality.

\subsubsection{3.1 San Jorge Gulf}\label{san-jorge-gulf}
